---
title: "R Notebook"
---

# Load libraries

```{r}

library(tidyverse)
library(magrittr)
library(lubridate)
library(ggalluvial)
library(ggrepel)
library(corrr) # for correlation matrix
library(patchwork) # for creating nice plot grids
library(colorspace) # better color palettes

```

# Theme options

```{r}

# Spruce up the default figure formatting
theme_set(theme_minimal() +
            theme(text = element_text(size = 10),
                  axis.title.x = element_text(size = 10,
                                              margin = margin(5, 0, 5, 0)),
                  axis.text.x = element_text(size = 8),
                  axis.title.y = element_text(size = 10,
                                              margin = margin(0, 5, 0, 5)),
                  axis.text.y = element_text(size = 8),
                  plot.title = element_text(size = 10),
                  panel.border = element_rect(color = "#E0E0E0", 
                                              size = 0.5, 
                                              fill = NA),
                  plot.margin = margin(5,5,5,5),
                  legend.key.size = unit(0.5, "cm"),
                  legend.text = element_text(size = 8)))

# Make sure axis labels (e.g. month names) are in English
Sys.setlocale("LC_ALL", "English")

# Create custom date scales for time-series plots
custom_week_scale <- function(start_date, end_date, format = "%d-%b") {
  scale_x_date(date_breaks = "1 week",
               date_minor_breaks = "1 day",
               date_labels = format,
               expand = c(0, 0),
               limits = c(ymd(start_date), ymd(end_date)))
}
custom_month_scale <- function(start_date, end_date, format = "%d-%b") {
  scale_x_date(date_breaks = "1 month",
               date_minor_breaks = "1 month",
               date_labels = format,
               expand = c(0, 0),
               limits = c(ymd(start_date), ymd(end_date)))
}

```

# Helper variables and functions

```{r}

analysis_start <- ymd("2020-01-01")
analysis_end <- ymd("2020-10-31")

assign_covid_preprint <- function(d) {
  d %>%
    mutate(covid_preprint = case_when(
      covid_preprint == T ~ "COVID-19 preprints",
      covid_preprint == F ~ "non-COVID-19 preprints"
    ))
}

# Function for basic cleaning of author names - convert to lowercase, remove whitespace, remove periods after initials
clean_names <- function(d) {
  d %>% 
    mutate(author_corresponding = str_to_lower(str_trim(str_remove(author_corresponding, "\\."))))
}

```


# Load datasets

```{r}

# Dataset of COVID-19 cases and deaths. Data is aggregated from the repository
# maintained by Johns Hopkins (https://github.com/CSSEGISandData/COVID-19)
covid <- read_csv("https://raw.githubusercontent.com/datasets/covid-19/master/data/worldwide-aggregate.csv")

# Full dataset of preprints from the launch of bioRxiv (Nov 2013) to the end of
# our study period (October 2020).
preprints_all <- read_csv("data/preprints_basic_20131101_20201031.csv")

# Abstracts of all preprints
#preprint_abstracts <- read_csv("data/preprint_abstracts_20131101_20201031.csv")

# More detailed preprint metadata covering period Sept 2019 - April 2020
preprints <- read_csv("data/preprints_full_20190101_20201031.csv",
                      col_types = cols(n_refs = "n", n_words = "n"))

# Preprint usage data (PDF downloads, abstract views)
preprint_usage <- read_csv("data/preprint_usage_20190101_20201031.csv")

# Preprint citation data
preprint_citations <- read_csv("data/preprint_citations_20190101_20201031.csv")

# Preprint altmetrics data
preprint_altmetrics <- read_csv("data/preprint_altmetrics_20190101_20201031.csv")

# Data of preprint timing (i.e. timing of first preprint per country - see Figure 2)
preprint_timing <- read_csv("data/preprint_earliest_per_country_20200101_20201031.csv")

# Preprint comments (counts) data
preprint_comments <- read_csv("data/preprint_comments_20190101_20201031.csv")

# Full dataset of preprints from other servers (note: dataset is generated
# externally and updated weekly)
preprints_other <- read_csv("https://raw.githubusercontent.com/nicholasmfraser/covid19_preprints/master/data/covid19_preprints.csv", col_types = "cccccc")

# Dataset of journal articles on COVID-19 from Dimensions
journal_articles <- read_csv("data/journal_articles_20200101_20201031.csv")

# Policy documents
#policy_docs <- read_csv("data/policy_docs.csv")

```

# Figure 1: Development of COVID-19, journal articles and preprints over time

```{r}

# Panel A: Development of COVID-19 cases and deaths

# Define  coordinates for annotations
x_1 <- ymd("2020-01-30")
x_2 <- ymd("2020-03-11")
y <- 50000000

p1A <- covid %>%
  filter(
    Date >= analysis_start,
    Date <= analysis_end) %>%
  select(Date, Confirmed, Deaths) %>%
  rename(Cases = Confirmed) %>%
  pivot_longer(Cases:Deaths) %>%
  ggplot(aes(x = Date, y = value, color = name)) +
  geom_line(size = 1) +
  geom_point(size = 1, shape = 21, fill = "white", stroke = 1) +
  geom_vline(aes(xintercept = x_1), size = 0.5, linetype = "dotted") +
  geom_vline(aes(xintercept = x_2), size = 0.5, linetype = "dotted") +
  geom_text(aes(x = x_1, y = y), label = " (i)", hjust = 0, color = "black") +
  geom_text(aes(x = x_2, y = y), label = " (ii)", hjust = 0, color = "black") +
  labs(x = "Date", y = "Confirmed Cases / Deaths", color = "") +
  scale_y_log10(labels = scales::comma) +
  custom_month_scale(analysis_start, analysis_end + 1) +
  scale_color_manual(values = qualitative_hcl(2, palette = "Dynamic")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggsave("outputs/figures/partials/Figure_1A.png", width = 10, height = 4)

# Panel B: Number of journal articles and preprints on COVID-19

p1B <- preprints_other %>%
  mutate(posted_date = ymd(posted_date), type = "Preprints") %>%
  bind_rows(journal_articles %>% mutate(posted_date = cr_created,
                                        type = "Journal Articles")) %>%
  count(type, posted_date) %>%
  complete(posted_date, nesting(type), fill = list(n = 0)) %>%
  group_by(type) %>%
  mutate(cumulative_n = cumsum(n),
         total_n = sum(n)) %>%
  ggplot() +
  geom_area(aes(x = posted_date, y = cumulative_n, fill = reorder(type, -total_n)), 
            color = "grey50", size = 0.5, alpha = 0.75) +
  labs(x = "Publication date", y = "Articles / Preprints", fill = "") +
  scale_y_continuous(labels = scales::comma) +
  custom_month_scale(analysis_start, analysis_end + 1) +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Cold")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggsave("outputs/figures/partials/Figure_1B.png", width = 10, height = 4)

# Panel C: Preprints on COVID-19 from multiple sources

# Find the repositories that have > X preprints
n_min <- 160

top_repositories <- preprints_other %>%
  filter(
    posted_date >= analysis_start,
    posted_date <= analysis_end) %>%
  count(source) %>%
  filter(n > n_min) %>%
  arrange(desc(n)) %>%
  pull(source)

fct_order <- c(top_repositories, "Other")

p1C <- preprints_other %>%
  mutate(posted_date = ymd(posted_date),
         source = case_when(
           source %in% top_repositories ~ source,
           T ~ "Other"),
         source = factor(source, levels = fct_order)) %>%
  count(source, posted_date) %>%
  complete(posted_date, nesting(source), fill = list(n = 0)) %>%
  group_by(source) %>%
  mutate(cumulative_n = cumsum(n),
         total_n = sum(n)) %>%
  ggplot() +
  geom_area(aes(x = posted_date, y = cumulative_n, fill = source), 
            color = "grey50", size = 0.5) +
  labs(x = "Publication date", y = "Preprints", fill = "") +
  scale_y_continuous(labels = scales::comma) +
  custom_month_scale(analysis_start, analysis_end + 1) +
  scale_fill_manual(values = qualitative_hcl(16, palette = "Set2")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggsave("outputs/figures/partials/Figure_1C.png", width = 10, height = 4)

# Patchwork
p1A + p1B + p1C + plot_layout(ncol = 1) +
  plot_annotation(tag_levels = "A") +
  ggsave("outputs/figures/Figure_1.png", width = 10, height = 10)

```

# Supplementary Figure 1: Number of preprints in relation to previous epidemics (Zika, Ebola)

```{r}

# Build search strings containing terms related to other epidemics
search_string_ebola <- "ebola|ZEBOV"
search_string_zika <- "zika|ZIKV"

ebola_start <- "2014-01-24"
ebola_end <- "2016-06-09"  # Outbreak declared ended: https://apps.who.int/iris/bitstream/handle/10665/208883/ebolasitrep_10Jun2016_eng.pdf

zika_start <- "2015-03-02"
zika_end <- "2016-11-18"  # Public health emergency declared ended: https://www.who.int/en/news-room/detail/18-11-2016-fifth-meeting-of-the-emergency-committee-under-the-international-health-regulations-(2005)-regarding-microcephaly-other-neurological-disorders-and-zika-virus

# Determine preprint counts of previous epidemics
covid_counts <- preprints_all %>% 
  filter(posted_date >= analysis_start & posted_date <= analysis_end) %>% 
  count(covid_preprint) %>%
  mutate(epidemic = "COVID-19") %>%
  rename(epi_preprint = covid_preprint)

ebola_counts <- preprints_all %>%
  inner_join(preprint_abstracts, by = "doi") %>%
  filter(posted_date >= ebola_start & posted_date <= ebola_end) %>%
  mutate(epi_preprint = case_when(
    str_detect(title, regex(search_string_ebola, ignore_case = TRUE)) ~ T,
    str_detect(abstract, regex(search_string_ebola, ignore_case = TRUE)) ~ T,
    T ~ F)) %>% 
  count(epi_preprint) %>%
  mutate(epidemic = "Western Africa Ebola virus")

zika_counts <- preprints_all %>% 
  inner_join(preprint_abstracts, by = "doi") %>%
  filter(posted_date >= zika_start & posted_date <= zika_end) %>%
  mutate(epi_preprint = case_when(
    str_detect(title, regex(search_string_zika, ignore_case = TRUE)) ~ T,
    str_detect(abstract, regex(search_string_ebola, ignore_case = TRUE)) ~ T,
    T ~ F)) %>% 
  count(epi_preprint) %>%
  mutate(epidemic = "Zika virus")

# Generate plot
pS1A <- bind_rows(covid_counts, ebola_counts, zika_counts) %>% 
  mutate(epi_preprint = case_when(
    epi_preprint == T ~ "Epidemic preprints",
    epi_preprint == F ~ "Non-epidemic preprints"),
    epidemic = factor(epidemic, 
                      levels = c("Western Africa Ebola virus", 
                                 "Zika virus", 
                                 "COVID-19"),
                      labels = c("Western Africa Ebola virus\n(2014/04/24 - 2016/06/09)", 
                                 "Zika virus\n(2015/03/02 - 2016/11/18)", 
                                 "COVID-19\n(2020/01/01 - 2020/10/31)"))) %>%
  ggplot(aes(x = epi_preprint, y = n, colour = epi_preprint, fill = epi_preprint)) +
  geom_col(color = "grey50", size = 0.25) +
  labs(x = "", y = "Total preprints", fill = "", color = "") +
  scale_color_manual(values = qualitative_hcl(2, palette = "Set2")) +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  facet_wrap(epidemic ~ ., strip.position = "bottom") +
  theme(axis.text.x = element_blank(),
        plot.margin = margin(0, 0, 20, 0)) +
  ggsave("outputs/figures/partials/Figure_S1A.png", width = 10, height = 4)

pS1B <-journal_articles %>% 
  count(article_is_oa, cr_created) %>%
  complete(cr_created, nesting(article_is_oa), fill = list(n = 0)) %>%
  group_by(article_is_oa) %>%
  mutate(cumulative_n = cumsum(n),
         total_n = sum(n)) %>%
  ungroup() %>%
  mutate(article_is_oa = case_when(
    article_is_oa == T ~ "Open Access",
    T  ~ "Closed Access")) %>%
  ggplot() +
  geom_area(aes(x = cr_created, y = cumulative_n, 
                fill = reorder(article_is_oa, -total_n)), 
            color = "grey50", size = 0.5) +
  labs(x = "Publication Date", y = "Published articles", fill = "") +
  scale_y_continuous(labels = scales::comma) +
  custom_month_scale("2020-01-01", "2020-11-01", format = "%d-%b") +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Dynamic")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggsave("outputs/figures/partials/Figure_S1B.png", width = 10, height = 4)

pS1A + pS1B + plot_layout(ncol = 1) +
  plot_annotation(tag_levels = "A") +
  ggsave("outputs/figures/Figure_S1.png", width = 10, height = 6)

```

# Figure 2: Preprint properties

```{r}

# Panel A: COVID vs non-COVID preprints for each server

p2A_properties <- preprints %>%
  filter(posted_date >= analysis_start,
         posted_date <= analysis_end) %>%
  assign_covid_preprint() %>%
  mutate(posted_month = floor_date(posted_date, unit = "month")) %>%
  count(covid_preprint, posted_month) %>%
  complete(posted_month, nesting(covid_preprint), fill = list(n = 0)) %>%
  ggplot(aes(x = posted_month, y = n, fill = covid_preprint)) +
  geom_col(color = "grey50", size = 0.25,
           position = position_dodge(preserve = "single")) +
  labs(x = "Posted month", y = "Preprints deposited", fill = "") +
  scale_y_continuous(labels = scales::comma) +
  scale_x_date(date_breaks = "1 month",
               date_minor_breaks = "1 month",
               date_labels = "%b") +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  ggsave("outputs/figures/partials/Figure_2A_properties.png", width = 6, height = 4)

p2B_properties <- preprints %>%
  filter(posted_date >= analysis_start,
         posted_date <= analysis_end) %>%
  count(covid_preprint, n_versions) %>%
  group_by(covid_preprint) %>%
  mutate(proportion = (n/sum(n)) * 100) %>%
  ungroup() %>%
  assign_covid_preprint() %>%
  ggplot(aes(x = n_versions, y = proportion, fill = covid_preprint)) +
  geom_col(color = "grey50", position = position_dodge(preserve = "single"), width = 0.75, size = 0.25) +
  labs(x = "Number of versions", y = "% of preprints", fill = "") +
  scale_x_continuous(breaks = seq(1:10)) +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  guides(fill = FALSE) +
  ggsave("outputs/figures/partials/Figure_2B_properties.png", width = 5, height = 4)

p2C_properties <- preprints %>%
  filter(posted_date >= analysis_start,
         posted_date <= analysis_end) %>%
  mutate(license = case_when(
    str_detect(license, "cc0") ~ "cc0",
    T ~ license
  )) %>%
  count(covid_preprint, license) %>%
  na.omit() %>%
  group_by(covid_preprint) %>%
  mutate(proportion = (n/sum(n)) * 100) %>%
  ungroup() %>%
  assign_covid_preprint() %>%
  mutate(license = factor(license,
                          levels = c("cc0", "cc_by", "cc_by_nc", "cc_by_nd", 
                                     "cc_by_nc_nd", "cc_no", "cc0_ng"),
                          labels = c("CC0", "CC BY", "CC BY-NC", "CC BY-ND", 
                                     "CC BY-NC-ND", "None", "CC0 NG?"))) %>%
  ggplot(aes(x = license, y = proportion, fill = covid_preprint)) +
  geom_col(color = "grey50", position = position_dodge(preserve = "single"), width = 0.75, size = 0.25) +
  labs(x = "License Type", y = "% of preprints", fill = "") +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  scale_x_discrete(label = function(x) stringr::str_wrap(x, 6)) +
  guides(fill = FALSE) +
  ggsave("outputs/figures/partials/Figure_2C_properties.png", width = 5, height = 4)

# Preprint screening time

p2D_properties <- preprints %>%
  filter(posted_date >= analysis_start,
         posted_date <= analysis_end) %>%
  mutate(doi_date = gsub("\\.", "-", substr(doi, 9, 18)),
         days = as.numeric(ymd(posted_date) - ymd(doi_date)),
         days = case_when(
           days > 7 ~ "8+",
           T ~ as.character(days)),
         days = factor(days,
                       levels = c("0", "1", "2", "3", "4", "5", "6", "7", "8+"),
                       labels = c("0", "1", "2", "3", "4", "5", "6", "7", "8+"))
  ) %>%
  count(covid_preprint, days) %>%
  group_by(covid_preprint) %>%
  mutate(proportion = (n/sum(n)) * 100) %>%
  ungroup() %>%
  assign_covid_preprint() %>%
  ggplot(aes(x = days, y = proportion,  fill = covid_preprint)) +
  geom_col(color = "grey50", position = position_dodge(preserve = "single"), 
           width = 0.75, size = 0.25) +    
  labs(x = "Screening time (days)", y = "% of preprints", fill = "") +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  guides(fill = FALSE) +
  ggsave("outputs/figures/partials/Figure_2D_properties.png", width = 5, height = 4)

p <- p2A_properties + p2B_properties + p2C_properties + p2D_properties + 
  plot_layout(ncol = 2, guides = 'collect') + 
  plot_annotation(tag_levels = "A") & theme(legend.position = 'bottom')

p + ggsave("outputs/figures/Figure_2_properties.png", width = 10, height = 6) 

```

# Supplementary Figure 2: Preprint properties

```{r}

# Panel A: COVID vs non-COVID preprints for each server

pS2A_properties <- preprints %>%
  filter(posted_date >= analysis_start,
         posted_date <= analysis_end) %>%
  assign_covid_preprint() %>%
  mutate(posted_month = floor_date(posted_date, unit = "month"),
         source = factor(source, 
                         levels = c("biorxiv", "medrxiv"),
                         labels = c("bioRxiv", "medRxiv"))) %>%
  count(covid_preprint, source, posted_month) %>%
  complete(posted_month, nesting(covid_preprint, source), fill = list(n = 0)) %>%
  ggplot(aes(x = posted_month, y = n, fill = covid_preprint)) +
  geom_col(color = "grey50", size = 0.25,
           position = position_dodge(preserve = "single")) +
  labs(x = "Posted month", y = "Preprints deposited", fill = "") +
  scale_y_continuous(labels = scales::comma) +
  scale_x_date(date_breaks = "1 month",
               date_minor_breaks = "1 month",
               date_labels = "%b") +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  facet_wrap(. ~ source, nrow = 2, strip.position = "right") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background = element_rect(color = "grey90", fill = "grey95"),
        legend.position = "top") +
  ggsave("outputs/figures/partials/Figure_S2A_properties.png", width = 6, height = 4)

pS2B_properties <- preprints %>%
  filter(posted_date >= analysis_start,
         posted_date <= analysis_end) %>%
  count(covid_preprint, source, n_versions) %>%
  group_by(covid_preprint, source) %>%
  mutate(proportion = (n/sum(n)) * 100) %>%
  ungroup() %>%
  assign_covid_preprint() %>%
  ggplot(aes(x = n_versions, y = proportion, fill = covid_preprint)) +
  geom_col(color = "grey50", position = position_dodge(preserve = "single"), width = 0.75, size = 0.25) +
  labs(x = "Number of versions", y = "% of preprints", fill = "") +
  scale_x_continuous(breaks = seq(1:10)) +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  guides(fill = FALSE) +
  facet_wrap(. ~ source, nrow = 2, strip.position = "right") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background = element_rect(color = "grey90", fill = "grey95"),
        legend.position = "top") +
  ggsave("outputs/figures/partials/Figure_S2B_properties.png", width = 5, height = 4)

pS2C_properties <- preprints %>%
  filter(posted_date >= analysis_start,
         posted_date <= analysis_end) %>%
  mutate(license = case_when(
    str_detect(license, "cc0") ~ "cc0",
    T ~ license
  )) %>%
  count(covid_preprint, source, license) %>%
  na.omit() %>%
  group_by(covid_preprint, source) %>%
  mutate(proportion = (n/sum(n)) * 100) %>%
  ungroup() %>%
  assign_covid_preprint() %>%
  mutate(license = factor(license,
                          levels = c("cc0", "cc_by", "cc_by_nc", "cc_by_nd", 
                                     "cc_by_nc_nd", "cc_no", "cc0_ng"),
                          labels = c("CC0", "CC BY", "CC BY-NC", "CC BY-ND", 
                                     "CC BY-NC-ND", "None", "CC0 NG?"))) %>%
  ggplot(aes(x = license, y = proportion, fill = covid_preprint)) +
  geom_col(color = "grey50", position = position_dodge(preserve = "single"), width = 0.75, size = 0.25) +
  labs(x = "License Type", y = "% of preprints", fill = "") +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  scale_x_discrete(label = function(x) stringr::str_wrap(x, 6)) +
  guides(fill = FALSE) +
  facet_wrap(. ~ source, nrow = 2, strip.position = "right") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background = element_rect(color = "grey90", fill = "grey95"),
        legend.position = "top") +
  ggsave("outputs/figures/partials/Figure_S2C_properties.png", width = 5, height = 4)

# Preprint screening time

pS2D_properties <- preprints %>%
  filter(posted_date >= analysis_start,
         posted_date <= analysis_end) %>%
  mutate(doi_date = gsub("\\.", "-", substr(doi, 9, 18)),
         days = as.numeric(ymd(posted_date) - ymd(doi_date)),
         days = case_when(
           days > 7 ~ "8+",
           T ~ as.character(days)),
         days = factor(days,
                       levels = c("0", "1", "2", "3", "4", "5", "6", "7", "8+"),
                       labels = c("0", "1", "2", "3", "4", "5", "6", "7", "8+"))
  ) %>%
  count(covid_preprint, source, days) %>%
  group_by(covid_preprint, source) %>%
  mutate(proportion = (n/sum(n)) * 100) %>%
  ungroup() %>%
  assign_covid_preprint() %>%
  ggplot(aes(x = days, y = proportion,  fill = covid_preprint)) +
  geom_col(color = "grey50", position = position_dodge(preserve = "single"), 
           width = 0.75, size = 0.25) +    
  labs(x = "Screening time (days)", y = "% of preprints", fill = "") +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  guides(fill = FALSE) +
  facet_wrap(. ~ source, nrow = 2, strip.position = "right") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background = element_rect(color = "grey90", fill = "grey95"),
        legend.position = "top") +
  ggsave("outputs/figures/partials/Figure_S2D_properties.png", width = 5, height = 4)

pS2 <- pS2A_properties + pS2B_properties + pS2C_properties + pS2D_properties + 
  plot_layout(ncol = 2, guides = 'collect') + 
  plot_annotation(tag_levels = "A") & theme(legend.position = 'bottom')

pS2 + ggsave("outputs/figures/Figure_S2_properties.png", width = 10, height = 6) 

```

# Figure 2: Authorships

```{r}

p2A <- preprints %>%
  filter(posted_date >= analysis_start,
         posted_date <= analysis_end) %>%
  mutate(n_authors = case_when(
    n_authors > 10 ~ "11+",
    T ~ as.character(n_authors)),
    n_authors = factor(n_authors,
                       levels = c(as.character(1:10), "11+"),
                       labels = c(as.character(1:10), "11+"))) %>%
  count(covid_preprint, n_authors) %>%
  group_by(covid_preprint) %>%
  mutate(proportion = (n/sum(n)) * 100) %>%
  ungroup() %>%
  assign_covid_preprint() %>%
  ggplot(aes(x = n_authors, y = proportion, fill = covid_preprint)) +
  geom_col(color = "grey50", position = position_dodge(preserve = "single"), 
           width = 0.75, size = 0.25) +
  labs(x = "Number of authors", y = "% of preprints", fill = "") +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  guides(fill = FALSE) +
  ggsave("outputs/figures/partials/Figure_2A_authors.png", width = 5, height = 4)

p2B <- preprints %>%
  filter(posted_date >= analysis_start,
         posted_date <= analysis_end,
         !is.na(institution_match_country_code)) %>%
  count(institution_match_country_code, covid_preprint) %>%
  mutate(continent = countrycode::countrycode(institution_match_country_code, 
                                              origin = 'iso2c', destination = 'continent')) %>% 
  pivot_wider(names_from = covid_preprint, values_from = n) %>%
  ggplot(aes(x = `TRUE`, y = `FALSE`, label = institution_match_country_code)) +
  geom_point(aes(color = continent), size = 1.5, alpha = 0.75) +
  geom_text_repel(force = 3,
                  segment.color = "grey50",
                  segment.size = 0.1,
                  segment.alpha = 0.75,
                  color = "grey25", 
                  size = 2) +
  labs(x = "COVID-19 preprints published",
       y = "non-COVID-19 preprints published",
       color = "") +
  scale_color_manual(values = qualitative_hcl(5, palette = "Set2")) +
  scale_y_log10() +
  scale_x_log10() +
  theme(legend.margin = margin(0, 0, 0, 0),
        legend.position = "top") +
  ggsave("outputs/figures/partials/Figure_2B_authors.png", width = 5, height = 4)

p2_authors <- p2A + p2B +
  plot_layout(ncol = 1, heights = c(1,2)) + 
  plot_annotation(tag_levels = "A")

p2_authors + ggsave("outputs/figures/Figure_2_authors.png", width = 5, height = 8) 

```

# Supplementary Figure 2: Authorships

```{r}
# Panel A: Science focus of senior author

# For each author, select their chronologically earliest COVID-19 preprint
preprint_author_covid <- preprints %>%
  clean_names() %>%
  filter(covid_preprint == T) %>% 
  group_by(author_corresponding) %>%
  arrange(posted_date) %>%
  slice(1) %>%
  select(author_corresponding, category, posted_date) %>%
  rename(category_covid = category,
         posted_covid = posted_date)

# For each author, select all their non-COVID-19 preprints
preprint_author_non_covid <- preprints_all %>%
  clean_names() %>%
  filter(covid_preprint == F) %>% 
  select(author_corresponding, category, posted_date) %>%
  rename(category_previous = category,
         posted_previous = posted_date)

# Filter to non-COVID-19 preprints uploaded before the latest COVID-19 preprint and select the chronologically latest
preprint_author_last <- preprint_author_covid %>%
  inner_join(preprint_author_non_covid, by = "author_corresponding") %>%
  filter(posted_covid >= posted_previous) %>%
  group_by(author_corresponding) %>%
  arrange(posted_previous) %>%
  slice(n()) %>%
  mutate(category_covid = tolower(category_covid),
         category_previous = tolower(category_previous))

total_cats <- c(preprint_author_last$category_covid, preprint_author_last$category_previous) %>% unique %>% sort

# Plot alluvial diagram
pS2A_authors <- preprint_author_last %>%
  mutate(category_covid = factor(category_covid, levels = total_cats),
         category_previous = factor(category_previous, levels = total_cats)) %>%
  group_by(category_covid, category_previous) %>%
  summarize(freq = n()) %>%
  ungroup() %>%
  filter(freq >= 10) %>%
  ggplot(aes(axis1 = category_previous, axis2 = category_covid, y = freq)) +
  scale_x_discrete(limits = c("Previous", "COVID-19"), 
                   expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Preprint categorisation") +
  geom_alluvium(aes(fill = category_previous), reverse = TRUE) +
  geom_stratum(reverse = TRUE, color = "grey50") + 
  geom_text(aes(label = after_stat(stratum)), size = 2.5, stat = "stratum", 
            reverse = TRUE, color = "grey25") +
  guides(fill = FALSE) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  ggsave("outputs/figures/partials/Figure_S2A_authors.png", width = 5, height = 8)

# Panel B: Country of authors by preprinter history

# Authors of COVID-19 preprints
covid_authors <- preprints %>% 
  filter(covid_preprint == T) %>%
  clean_names() %>%
  pull(author_corresponding)

# Control authors (authors not in COVID-19 group)
control_authors <- preprints %>% 
  filter(covid_preprint == F & !(author_corresponding %in% covid_authors)) %>%
  clean_names() %>%
  pull(author_corresponding)

# Determine preprinter history
preprinter_history <- preprints_all %>%
  filter(posted_date < analysis_start,
         author_corresponding != "? ?",
         !is.na(author_corresponding)) %>%
  clean_names() %>%
  mutate(author_corresponding = factor(author_corresponding, 
                                       levels = unique(c(covid_authors, control_authors)))) %>%
  group_by(author_corresponding, .drop = F) %>%
  summarize(n = n()) %>%
  mutate(
    author_group = case_when(
      author_corresponding %in% covid_authors ~ "COVID-19 preprints",
      author_corresponding %in% control_authors ~ "non-COVID-19 preprints"
    ),
    preprinter_status = case_when(
      n == 0 ~ "First-time", 
      T ~ "Previous")
  ) %>%
  distinct(author_corresponding, author_group, preprinter_status) %>%
  filter(author_corresponding %in% (preprints %>% 
                                      filter(posted_date >= analysis_start,
                                             posted_date <= analysis_end) %>% 
                                      clean_names() %>%
                                      pull(author_corresponding))) # Limit to authors during the study period only

# Append author locations and calculate proportions for plot
preprinter_history_proportions <- preprints %>%
  clean_names() %>%
  inner_join(preprinter_history, by = "author_corresponding") %>%
  distinct(author_corresponding, author_group, preprinter_status, institution_match_country_name) %>%
  count(author_group, preprinter_status, institution_match_country_name) %>% 
  complete(author_group, preprinter_status, institution_match_country_name, 
           fill = list(n = 0)) %>%
  group_by(author_group, preprinter_status) %>%
  mutate(prop = n*100 / sum(n))

countries_to_plot <- preprinter_history_plot %>% 
  filter(author_group == "non-COVID-19 preprints" & preprinter_status == "Previous") %>% 
  arrange(-prop) %>% slice(1:15) %>% 
  pull(institution_match_country_name)

pS2B_authors <- preprinter_history_proportions %>%
  filter(institution_match_country_name %in% countries_to_plot) %>%
  ggplot() +
  geom_bar(mapping = aes(x = institution_match_country_name, y = prop, 
                         fill = author_group, alpha = preprinter_status), 
           stat = "identity", position = "dodge", color = "grey50", size = 0.25) + 
  coord_flip() +
  scale_x_discrete(limits = rev(countries_to_plot)) +
  scale_y_continuous(expand = c(0,0)) +
  facet_wrap(facets = ~author_group) +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  scale_alpha_manual(values = c(1.0, 0.4),
                     breaks = c("Previous", "First-time")) +
  labs(x = "Country", y = "% of senior authors") +
  guides(fill = FALSE) +
  theme(legend.title = element_blank(),
        legend.margin=margin(c(0,5,5,5)),
        legend.position = c(0.8, 0.12),
        strip.text = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.background = element_rect(fill = "white", color = "grey50")) +
  ggsave("outputs/figures/partials/Figure_S2B_authors.png", width = 8, height = 5)

# Panel C: Time between first case and first preprint

pS2C_authors <- preprint_timing %>%
  mutate(continent = countrycode::countrycode(institution_match_country_code, 
                                              origin = 'iso2c', destination = 'continent')) %>% 
  filter(!is.na(first_preprint) & !is.na(first_case) & !is.na(continent)) %>%
  filter(first_case <= "2020-04-30",
         first_preprint <= "2020-04-30") %>%
  ggplot(aes(x = first_case, y = first_preprint, label = institution_match_country_code)) +
  geom_point(aes(color = continent), size=2) +
  geom_text_repel(force = 5,
                  segment.color = "grey50",
                  segment.size = 0.2,
                  segment.alpha = 0.8,
                  color = "grey25", 
                  size = 2.5) +
  geom_abline(intercept = 0, slope = 1, color = "grey50", lty = "dashed", alpha = 0.5) +
  labs(x = "Date of first case",
       y = "Date of first preprint",
       color = "") +
  scale_color_manual(values = qualitative_hcl(5, palette = "Set2")) +
  coord_cartesian(ylim = c(ymd("2020-01-01", "2020-04-30")),
                  xlim = c(ymd("2020-01-01", "2020-04-30"))) +
  theme(legend.margin = margin(0, 0, 0, 0),
        legend.position = "top") +
  ggsave("outputs/figures/partials/Figure_S2C_authors.png", width = 5, height = 4)


layout <- "
AABB
AACC
"
p <- pS2A_authors + pS2B_authors + pS2C_authors +
  plot_layout(design = layout) & theme(plot.margin = margin(r = 10))

p + plot_annotation(tag_levels = "A") +
  ggsave("outputs/figures/Figure_S2_authors.png", width = 10, height = 8) 

```

# Figure 2: Publication outcomes

```{r}

# Panel A: proportions of preprints published over time

p2A_outcomes <- preprints %>% 
  filter(posted_date >= analysis_start,
         posted_date <= analysis_end) %>%
  assign_covid_preprint() %>%
  mutate(posted_month = floor_date(posted_date, unit = "month"),
         is_published = !is.na(published_doi)) %>%
  count(posted_month, covid_preprint, is_published) %>%
  group_by(posted_month, covid_preprint) %>%
  mutate(prop = n*100 / sum(n)) %>%
  ungroup() %>%
  filter(is_published == T) %>%
  ggplot(aes(x = posted_month, y = prop, fill = covid_preprint)) +
  geom_col(color = "grey50", size = 0.25, position = "dodge") +
  labs(x = "", y = "% of preprints published", fill ="") +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  scale_x_date(date_breaks = "1 month",
               date_minor_breaks = "1 month",
               date_labels = "%b") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "top") +
  ggsave("outputs/figures/partials/Figure_2A_outcomes.png", width = 5, height = 4)

# Panel B: time to publication

p2B_outcomes <- preprints %>%
  mutate(covid_preprint = case_when(
    (covid_preprint == T & posted_date >= analysis_start & posted_date <= analysis_end) ~ "COVID-19 preprints",
    (covid_preprint == F & posted_date >= analysis_start & posted_date <= analysis_end) ~ "non-COVID-19 preprints",
    T ~ "preprints 2019")) %>%
  mutate(pub_bracket = cut(as.numeric(delay_in_days), 
                           seq(0, 360, by = 10), 
                           labels=seq(0, 350, by = 10))) %>%
  group_by(covid_preprint) %>%
  count(pub_bracket) %>%
  mutate(prop = n*100 / sum(n)) %>%
  filter(!is.na(pub_bracket)) %>%
  ggplot(aes(x = pub_bracket, y = prop, fill=covid_preprint, color=covid_preprint)) +
  geom_bar(alpha = 0.25, width = 1, size = 0.25, 
           stat = "identity", position="identity") +
  scale_x_discrete(breaks = as.character(seq(0, 350, by = 50))) +
  labs(x = "Time from preprint posting to publication (days)",
       y = "% of all preprints") +
  scale_color_manual(values = c(qualitative_hcl(2, palette = "Set2"), "grey60")) +
  scale_fill_manual(values = c(qualitative_hcl(2, palette = "Set2"), "grey60")) +
  theme(legend.title = element_blank(),
        legend.position = "top",
        legend.text = element_text(size = 8)) +
  ggsave("outputs/figures/partials/Figure_2B_outcomes.png", width = 5, height = 4)

# Panel C: Journal locations of published papers (top 10)

p2C_outcomes <- preprints %>%
  filter(covid_preprint == T) %>%
  count(published_journal) %>%
  na.omit() %>%
  arrange(-n) %>%
  slice(1:10) %>%
  ggplot(aes(x = reorder(published_journal, -n), y = n)) +
  geom_col(color = "grey50", fill = qualitative_hcl(n = 1, palette = "Set2"),
           position = "dodge", size = 0.25) +
  labs(x = "", y = str_wrap("Number of COVID-19 preprints published", 20), fill = "") +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 25)) +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1)) +
  ggsave("outputs/figures/partials/Figure_2C_outcomes.png", width = 5, height = 4) 

# Panel D: Time to publication by publisher

top_publishers <- preprints %>%
  filter(covid_preprint == T,
         posted_date >= analysis_start,
         posted_date <= analysis_end) %>%
  count(published_publisher) %>%
  na.omit() %>%
  top_n(10, n) %>%
  pull(published_publisher)

p2D_outcomes <- preprints %>%
  filter(posted_date >= analysis_start,
         posted_date <= analysis_end,
         published_publisher %in% top_publishers) %>%
  assign_covid_preprint() %>%
  mutate(published_publisher = factor(published_publisher,
                                      levels = c("American Association for the Advancement of Science (AAAS)",
                                                 "American Society for Microbiology",
                                                 "BMJ",
                                                 "Elsevier BV",
                                                 "Frontiers Media SA",
                                                 "MDPI AG",
                                                 "Oxford University Press (OUP)",
                                                 "Public Library of Science (PLoS)",
                                                 "Springer Science and Business Media LLC",
                                                 "Wiley"),
                                      labels = c("AAAS", "ASM", "BMJ", "Elsevier", "Frontiers", 
                                                 "MDPI", "OUP", "PLoS", "Springer", "Wiley"))) %>%
  filter(delay_in_days > 0) %>%
  ggplot(aes(x = published_publisher, y = delay_in_days, color = covid_preprint)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(fill = factor(covid_preprint)),
              shape = 21, size = 0.2, alpha = 0.2,
              position = position_jitterdodge(jitter.width = 0.5)) + 
  labs(x = "", y = str_wrap("Time from preprint posting to publication (days)", 30), 
       fill = "", color = "") +
  scale_color_manual(values = qualitative_hcl(2, palette = "Set2")) +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
        legend.position = "top") +
  guides(fill = FALSE, colour = FALSE) +
  ggsave("outputs/figures/partials/Figure_2D_outcomes.png", width = 5, height = 4) 


p <- p2A_outcomes + p2B_outcomes + p2C_outcomes + p2D_outcomes +
  plot_layout(ncol = 2)

p + plot_annotation(tag_levels = "A") +
  ggsave("outputs/figures/Figure_2_outcomes.png", width = 10, height = 8) 

```

---- OLD FIGURES BELOW ------

# Figure 2
```{r}

# Panel A: Numbers of COVID vs non-COVID preprints
# Limit to preprints posted between January and April 2020

p2A <- preprints %>%
  filter(posted_date >= "2020-01-01") %>%
  mutate(posted_week = floor_date(posted_date, unit = "week", 
                                  week_start = 1)) %>%
  assign_covid_preprint() %>%
  count(covid_preprint, posted_week) %>%
  complete(posted_week, nesting(covid_preprint), fill = list(n = 0)) %>%
  ggplot(aes(x = posted_week, y = n, fill = covid_preprint)) +
  geom_col(color = "grey50", position = position_dodge(preserve = "single"), 
           width = 6, size = 0.25) +
  labs(x = "Posted date (week beginning)", y = "Preprints deposited", fill = "") +
  scale_y_continuous(labels = scales::comma) +
  custom_date_scale("2020-01-01", NA) +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "top") +
  ggsave("outputs/figures/partials/Figure_2A.png", width = 10, height = 4)

# Panel B: Preprint screening time

p2B <- preprints %>%
  filter(posted_date >= "2020-01-01")  %>%
  mutate(doi_date = gsub("\\.", "-", substr(doi, 9, 18)),
         days = as.numeric(ymd(posted_date) - ymd(doi_date)),
         days = case_when(
           days > 7 ~ "8+",
           T ~ as.character(days)),
         days = factor(days,
                       levels = c("0", "1", "2", "3", "4", "5", "6", "7", "8+"),
                       labels = c("0", "1", "2", "3", "4", "5", "6", "7", "8+"))
  ) %>%
  count(covid_preprint, days) %>%
  group_by(covid_preprint) %>%
  mutate(proportion = (n/sum(n)) * 100) %>%
  ungroup() %>%
  assign_covid_preprint() %>%
  ggplot(aes(x = days, y = proportion,  fill = covid_preprint)) +
  geom_col(color = "grey50", position = position_dodge(preserve = "single"), 
           width = 0.75, size = 0.25) +    
  labs(x = "Screening time (days)", y = "% of preprints", fill = "") +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  guides(fill = FALSE) +
  ggsave("outputs/figures/partials/Figure_2B.png", width = 5, height = 4)

# Panel C: Author counts

p2C <- preprints %>%
  filter(posted_date >= "2020-01-01") %>%
  mutate(n_authors = map_dbl(authors, ~(length(str_split(., ";")[[1]]))),
         n_authors = case_when(
           n_authors > 10 ~ "11+",
           T ~ as.character(n_authors)),
         n_authors = factor(n_authors,
                            levels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11+"),
                            labels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11+"))
  ) %>%
  count(covid_preprint, n_authors) %>%
  group_by(covid_preprint) %>%
  mutate(proportion = (n/sum(n)) * 100) %>%
  ungroup() %>%
  assign_covid_preprint() %>%
  ggplot(aes(x = n_authors, y = proportion, fill = covid_preprint)) +
  geom_col(color = "grey50", position = position_dodge(preserve = "single"), 
           width = 0.75, size = 0.25) +
  labs(x = "Number of authors", y = "% of preprints", fill = "") +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  guides(fill = FALSE) +
  ggsave("outputs/figures/partials/Figure_2C.png", width = 5, height = 4)

# Panel D: Author attributes: country, nationality (top 10 countries)
# Extract and classify all senior authors
covid_authors <- preprints %>% 
  filter(covid_preprint == T) %>%
  clean_names() %>%
  pull(author_corresponding)

control_authors <- preprints %>% 
  filter(covid_preprint == F & !(author_corresponding %in% covid_authors)) %>%
  clean_names() %>%
  pull(author_corresponding)

preprinter_history <- preprints_all %>%
  filter(posted_date < "2020-01-01") %>%
  clean_names() %>%
  mutate(author_corresponding = factor(author_corresponding, 
                                       levels = unique(c(covid_authors, control_authors)))) %>%
  filter(!is.na(author_corresponding)) %>%
  group_by(author_corresponding, .drop = FALSE) %>%
  summarise(n = n()) %>%
  mutate(
    author_group = case_when(
      author_corresponding %in% covid_authors ~ "COVID-19 preprints",
      author_corresponding %in% control_authors ~ "non-COVID-19 preprints"
    ),
    preprinter_status = case_when(
      n == 0 ~ "First-time", 
      T ~ "Previous")
  )

# Append author locations and calculate proportions for plot
preprinter_history_plot <- preprinter_history %>%
  left_join(preprints %>%
              select(author_corresponding, institution_match_country_name) %>%
              distinct() %>% 
              clean_names(), 
            by = "author_corresponding") %>%
  filter(!is.na(institution_match_country_name),
         author_corresponding != "? ?") %>%
  select(-n) %>%
  count(author_group, preprinter_status, institution_match_country_name) %>% 
  complete(author_group, preprinter_status, institution_match_country_name, 
           fill = list(n = 0)) %>%
  group_by(author_group, preprinter_status) %>%
  mutate(prop = n*100 / sum(n))

countries_to_plot <- preprinter_history_plot %>% 
  filter(author_group == "non-COVID-19 preprints" & preprinter_status == "Previous") %>% 
  arrange(-prop) %>% slice(1:15) %>% 
  pull(institution_match_country_name)

p2D <- preprinter_history_plot %>%
  filter(institution_match_country_name %in% countries_to_plot) %>%
  ggplot() +
  geom_bar(mapping = aes(x = institution_match_country_name, y = prop, 
                         fill = author_group, alpha = preprinter_status), 
           stat = "identity", position = "dodge", color = "grey50", size = 0.25) + 
  coord_flip() +
  scale_x_discrete(limits = rev(countries_to_plot)) +
  scale_y_continuous(expand = c(0,0)) +
  facet_wrap(facets = ~author_group) +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  scale_alpha_manual(values = c(1.0, 0.4),
                     breaks = c("Previous", "First-time")) +
  labs(x = "Country", y = "% of senior authors") +
  guides(fill = FALSE) +
  theme(legend.title = element_blank(),
        legend.margin=margin(c(0,5,5,5)),
        legend.position = c(0.8, 0.12),
        strip.text = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.background = element_rect(fill = "white", color = "grey50")) +
  ggsave("outputs/figures/partials/Figure_2D.png", width = 8, height = 5)

# Panel E: License types

p2E <- preprints %>%
  filter(posted_date >= "2020-01-01") %>%
  mutate(license = case_when(
    str_detect(license, "cc0") ~ "cc0",
    T ~ license
  )) %>%
  count(covid_preprint, license) %>%
  na.omit() %>%
  group_by(covid_preprint) %>%
  mutate(proportion = (n/sum(n)) * 100) %>%
  ungroup() %>%
  assign_covid_preprint() %>%
  mutate(license = factor(license,
                          levels = c("cc0", "cc_by", "cc_by_nc", "cc_by_nd", 
                                     "cc_by_nc_nd", "cc_no", "cc0_ng"),
                          labels = c("CC0", "CC BY", "CC BY-NC", "CC BY-ND", 
                                     "CC BY-NC-ND", "None", "CC0 NG?"))) %>%
  ggplot(aes(x = license, y = proportion, fill = covid_preprint)) +
  geom_col(color = "grey50", position = position_dodge(preserve = "single"), width = 0.75, size = 0.25) +
  labs(x = "License Type", y = "% of preprints", fill = "") +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  scale_x_discrete(label = function(x) stringr::str_wrap(x, 6)) +
  guides(fill = FALSE) +
  ggsave("outputs/figures/partials/Figure_2E.png", width = 5, height = 4)

# Panel F: Revisions per preprint for COVID vs non-COVID

p2F <- preprints %>%
  filter(posted_date >= "2020-01-01") %>%
  count(covid_preprint, n_versions) %>%
  group_by(covid_preprint) %>%
  mutate(proportion = (n/sum(n)) * 100) %>%
  ungroup() %>%
  assign_covid_preprint() %>%
  ggplot(aes(x = n_versions, y = proportion, fill = covid_preprint)) +
  geom_col(color = "grey50", position = position_dodge(preserve = "single"), width = 0.75, size = 0.25) +
  labs(x = "Number of versions", y = "% of preprints", fill = "") +
  scale_x_continuous(breaks = seq(1:10)) +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  guides(fill = FALSE) +
  ggsave("outputs/figures/partials/Figure_2F.png", width = 5, height = 4)

# Panel G: Word counts

p2G <- preprints %>%
  filter(posted_date >= "2020-01-01",
         source == "biorxiv",
         n_words != 0) %>%
  assign_covid_preprint() %>%
  ggplot(aes(x = covid_preprint, y = n_words, color = covid_preprint)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(shape = 21, size = 0.2, alpha = 0.2, width = 0.3) +
  labs(x = "", y = "Word Count", fill = "") +
  scale_color_manual(values = qualitative_hcl(2, palette = "Set2")) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 12)) +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(size = 7.5)) +
  ggsave("outputs/figures/partials/Figure_2G.png", width = 5, height = 4)

# Panel H: Reference counts

p2H <- preprints %>%
  filter(posted_date >= "2020-01-01",
         source == "biorxiv") %>%
  assign_covid_preprint() %>%
  ggplot(aes(x = covid_preprint, y = n_refs, color = covid_preprint)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(shape = 21, size = 0.2, alpha = 0.2, width = 0.3) +
  labs(x = "", y = "Reference Count", fill = "") +
  scale_color_manual(values = qualitative_hcl(2, palette = "Set2")) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 12)) +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(size = 7.5)) +
  ggsave("outputs/figures/partials/Figure_2H.png", width = 5, height = 4)

# Panel I: Percentage of preprints published

p2I <- preprints %>% 
  filter(posted_date >= "2020-01-01") %>%
  assign_covid_preprint() %>%
  count(covid_preprint, is_published) %>%
  group_by(covid_preprint) %>%
  mutate(prop = n*100 / sum(n)) %>%
  filter(is_published == T) %>%
  ggplot(aes(x = covid_preprint, y = prop, fill=covid_preprint)) +
  geom_col(color = "grey50", size = 0.25, position = "dodge") +
  labs(x = "", y = "% of preprints published", fill ="") +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 12)) +
  theme(axis.title.x = element_blank()) +
  guides(fill = FALSE) + 
  theme(axis.text.x = element_text(size = 7.5)) +
  ggsave("outputs/figures/partials/Figure_2I.png", width = 5, height = 4)

# Panel J: Publishing timeline

p2J <- preprints %>%
  mutate(covid_preprint = case_when(
    covid_preprint == T ~ "COVID-19 preprints",
    (covid_preprint == F & posted_date >= "2020-01-01") ~ "non-COVID-19 preprints",
    T ~ "Preprints Sep-Dec 2019")) %>%
  mutate(pub_bracket = cut(as.numeric(delay_in_days), 
                           seq(0, 240, by = 10), 
                           labels=seq(0, 230, by = 10))) %>%
  group_by(covid_preprint) %>%
  count(pub_bracket) %>%
  mutate(prop = n*100 / sum(n)) %>%
  filter(!is.na(pub_bracket)) %>%
  ggplot(aes(x = pub_bracket, y = prop, fill=covid_preprint, color=covid_preprint)) +
  geom_bar(alpha = 0.25, width = 1, size = 0.25, 
           stat = "identity", position="identity") +
  scale_x_discrete(labels = c("0","","","","","50","","","","","100","","","","","150","","","","","200","","","")) +
  labs(x = "Time from preprint posting to publication (days)",
       y = "% of all preprints") +
  scale_color_manual(values = c(qualitative_hcl(2, palette = "Set2"), "grey60")) +
  scale_fill_manual(values = c(qualitative_hcl(2, palette = "Set2"), "grey60")) +
  theme(legend.title = element_blank(),
        legend.position = "top",
        legend.text = element_text(size = 8)) +
  ggsave("outputs/figures/partials/Figure_2J.png", width = 5, height = 4)

# define layout manually - see https://patchwork.data-imaginist.com/articles/guides/layout.html
layout <- "
AAAA
BBCC
DDEE
DDFF
DDGH
IJJJ
"
p <- p2A + p2B + p2C + p2D + p2E + p2F + p2G + p2H + p2I + p2J +
  plot_layout(design = layout)

p + plot_annotation(tag_levels = "A") +
  ggsave("outputs/figures/Figure_2.png", width = 10, height = 12) 

```

# Supplementary Figure 2:

```{r}

# Panel A: COVID vs non-COVID preprints for bioRxiv only

pS2A <- preprints %>%
  filter(posted_date >= "2020-01-01") %>%
  assign_covid_preprint() %>%
  mutate(posted_week = floor_date(posted_date, unit = "week", week_start = 1),
         source = factor(source, 
                         levels = c("biorxiv", "medrxiv"),
                         labels = c("bioRxiv", "medRxiv"))) %>%
  count(covid_preprint, source, posted_week) %>%
  complete(posted_week, nesting(covid_preprint, source), fill = list(n = 0)) %>%
  ggplot(aes(x = posted_week, y = n, fill = covid_preprint)) +
  geom_col(color = "grey50", size = 0.25,
           position = position_dodge(preserve = "single")) +
  labs(x = "Posted date (week beginning)", y = "Preprints deposited", fill = "") +
  scale_y_continuous(labels = scales::comma) +
  custom_date_scale("2020-01-01", NA) +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  facet_wrap(. ~ source, nrow = 2, strip.position = "right") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background = element_rect(color = "grey90", fill = "grey95"),
        legend.position = "top") +
  ggsave("outputs/figures/partials/Figure_S2A.png", width = 10, height = 4)

# Panel B: Preprint screening time by server

pS2B <- preprints %>%
  filter(posted_date >= "2020-01-01")  %>%
  mutate(doi_date = gsub("\\.", "-", substr(doi, 9, 18)),
         days = as.numeric(ymd(posted_date) - ymd(doi_date))) %>%
  count(covid_preprint, source, days) %>%
  group_by(covid_preprint, source) %>%
  mutate(proportion = (n/sum(n)) * 100) %>%
  ungroup() %>%
  mutate(covid_preprint = case_when(
    covid_preprint == T ~ "COVID-19 preprints",
    covid_preprint == F ~ "non-COVID-19 preprints"
  ),
  source = factor(source, 
                  levels = c("biorxiv", "medrxiv"),
                  labels = c("bioRxiv", "medRxiv"))) %>%
  ggplot(aes(x = days, y = proportion,  fill = covid_preprint)) +
  geom_col(color = "grey50", size = 0.25,
           position = position_dodge(preserve = "single")) +    
  facet_wrap(. ~ source, nrow = 2, strip.position = "right") +
  labs(x = "Screening time (days)", y = "% of preprints", fill = "") +
  scale_x_continuous(breaks = seq(from = 0, to = 14, by = 1)) +
  coord_cartesian(xlim = c(0, 14)) +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  guides(fill = FALSE) +
  theme(strip.background = element_rect(color = "grey90", fill = "grey95")) +
  ggsave("outputs/figures/partials/Figure_S2B.png", width = 10, height = 4)

# Panel C: Science focus of senior author
# For each author, select their chronologically latest COVID-19 preprint
preprint_author_last_covid <- preprints %>%
  clean_names() %>%
  filter(covid_preprint == T) %>% 
  arrange(posted_date) %>% 
  group_by(author_corresponding) %>% 
  slice(n()) %>%
  select(author_corresponding, category, posted_date) %>%
  rename(category_covid = category,
         posted_covid = posted_date)

# For each author, select their non-COVID-19 preprints
preprint_author_last_non_covid <- preprints_all %>%
  clean_names() %>%
  filter(covid_preprint == F) %>% 
  arrange(posted_date) %>% 
  group_by(author_corresponding) %>% 
  select(author_corresponding, category, posted_date) %>%
  rename(category_previous = category,
         posted_previous = posted_date)

# Filter to non-COVID-19 preprints uploaded before the latest COVID-19 preprint and select the chronologically latest
preprint_author_last <- preprint_author_last_covid %>%
  inner_join(preprint_author_last_non_covid, by = "author_corresponding") %>%
  filter(posted_covid >= posted_previous) %>%
  slice(n()) %>%
  mutate(category_covid = tolower(category_covid),
         category_previous = tolower(category_previous))

total_cats <- c(preprint_author_last$category_covid, preprint_author_last$category_previous) %>% unique %>% sort

# Plot alluvial diagram
pS2C <- preprint_author_last %>%
  mutate(category_covid = factor(category_covid, levels = total_cats),
         category_previous = factor(category_previous, levels = total_cats)) %>%
  group_by(category_covid, category_previous) %>%
  summarize(freq = n()) %>%
  ungroup() %>%
  filter(freq >= 5) %>%
  ggplot(aes(axis1 = category_previous, axis2 = category_covid, y = freq)) +
  scale_x_discrete(limits = c("Previous", "COVID-19"), 
                   expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Preprint categorisation") +
  geom_alluvium(aes(fill = category_previous), reverse = TRUE) +
  geom_stratum(reverse = TRUE, color = "grey50") + 
  geom_text(aes(label = after_stat(stratum)), size = 2.5, stat = "stratum", 
            reverse = TRUE, color = "grey25") +
  guides(fill = FALSE) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  ggsave("outputs/figures/partials/Figure_S2D.png", width = 5, height = 8)

# Panel D: First case -> first preprint by location

pS2D <- preprint_timing %>%
  mutate(continent = countrycode::countrycode(institution_match_country_code, 
                                              origin = 'iso2c', destination = 'continent')) %>% 
  filter(!is.na(first_preprint) & !is.na(first_case) & !is.na(continent)) %>%
  ggplot(aes(x = first_case, y = first_preprint, label = institution_match_country_code)) +
  geom_point(aes(color = continent), size=2) +
  geom_text_repel(force = 5,
                  segment.color = "grey50",
                  segment.size = 0.2,
                  segment.alpha = 0.8,
                  color = "grey25", 
                  size = 2.5) +
  geom_abline(intercept = 0, slope = 1, color = "grey50", lty = "dashed", alpha = 0.5) +
  labs(x = "Date of first case",
       y = "Date of first preprint",
       color = "") +
  scale_color_manual(values = qualitative_hcl(5, palette = "Set2")) +
  coord_cartesian(ylim = c(ymd("2020-01-01", "2020-04-30")),
                  xlim = c(ymd("2020-01-01", "2020-04-30"))) +
  theme(legend.margin = margin(0, 0, 0, 0),
        legend.position = "top") +
  ggsave("outputs/figures/partials/Figure_S2C.png", width = 5, height = 4)



# Panel E: Journal locations of published papers (top 10)

pS2E <- preprints %>%
  filter(covid_preprint == T) %>%
  count(published_journal) %>%
  na.omit() %>%
  arrange(-n) %>%
  slice(1:10) %>%
  ggplot(aes(x = published_journal, y = n)) +
  geom_col(color = "grey50", fill = qualitative_hcl(n = 1, palette = "Set2"),
           position = "dodge", size = 0.25) +
  labs(x = "", y = str_wrap("Number of COVID-19 preprints published", 20), fill = "") +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 25)) +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1)) +
  ggsave("outputs/figures/partials/Figure_S2E.png", width = 5, height = 4) 

# Panel F: Time to publication for different publishers (top 8)

top_publishers <- preprints %>%
  filter(covid_preprint == T,
         posted_date >= "2020-01-01") %>%
  count(published_publisher) %>%
  na.omit() %>%
  arrange(-n) %>%
  slice(1:8) %>%
  pull(published_publisher)

pS2F <- preprints %>%
  filter(posted_date >= "2020-01-01",
         published_publisher %in% top_publishers) %>%
  assign_covid_preprint() %>%
  mutate(published_publisher = factor(published_publisher,
                                      levels = c("AME Publishing Company",
                                                 "American Association for the Advancement of Science (AAAS)",
                                                 "American Society for Microbiology",
                                                 "Elsevier BV",
                                                 "Frontiers Media SA",
                                                 "MDPI AG",
                                                 "Oxford University Press (OUP)",
                                                 "Public Library of Science (PLoS)",
                                                 "Springer Science and Business Media LLC",
                                                 "Wiley"),
                                      labels = c("AME", "AAAS", "ASM", "Elsevier", "Frontiers", 
                                                 "MDPI", "OUP", "PLoS", "Springer", "Wiley"))) %>%
  filter(delay_in_days > 0) %>%
  ggplot(aes(x = published_publisher, y = delay_in_days, color = covid_preprint)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(fill = factor(covid_preprint)),
              shape = 21, size = 0.2, alpha = 0.2,
              position = position_jitterdodge(jitter.width = 0.5)) + 
  labs(x = "", y = str_wrap("Time from preprint posting to publication (days)", 30), 
       fill = "", color = "") +
  scale_color_manual(values = qualitative_hcl(2, palette = "Set2")) +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
        legend.position = "top") +
  guides(fill = FALSE, colour = FALSE) +
  ggsave("outputs/figures/partials/Figure_S2F.png", width = 5, height = 4) 


# Panel G: Acceptance rates of COVID, non-COVID works
# Reconstruct aggregated totals from confidential data supplied by publishers
accept_rates <- structure(list(percentage = c(15.92079208, 26.51880844), 
                               is_covid = c("COVID-19 manuscripts", "non-COVID-19 manuscripts")), 
                          class = "data.frame", row.names = c(NA, -2L))

pS2G <- accept_rates %>%
  ggplot(aes(x = is_covid, y = percentage, fill=is_covid)) +
  geom_col(color = "grey50", size = 0.25, position = "dodge") +
  labs(x = "", y = "% accepted", fill ="") +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  theme(axis.title.x = element_blank()) +
  guides(fill = FALSE) +
  ggsave("outputs/figures/partials/Figure_S2G.png", width = 5, height = 4)

# Patchwork

# define layout 
layout <- "
AB
CD
CD
CE
FG
"

p <- pS2A + pS2B + pS2C + pS2D + pS2E + pS2F + pS2G +
  plot_layout(design = layout, heights = c(1.5, 1, 1, 1, 1)) & theme(legend.position = "top")

p + plot_annotation(tag_levels = "A") +
  ggsave("outputs/figures/Figure_S2.png", width = 10, height = 12)

```

# Figure 3: Preprint access

```{r}

# Prepare data
d <- preprint_usage %>%
  inner_join(preprints, by = "doi") %>%
  group_by(doi, posted_date, covid_preprint) %>%
  summarize(full_text_views = sum(full_text_views),
            abstract_views = sum(abstract_views),
            pdf_downloads = sum(pdf_downloads)) %>%
  ungroup() %>%
  assign_covid_preprint() %>%
  mutate(posted_month = floor_date(posted_date, unit = "month")) %>%
  filter(posted_month >= analysis_start,
         posted_month <= analysis_end)

# Abstract views
p3A <- d %>%
  ggplot(aes(x = posted_month, y = abstract_views, color = covid_preprint,
             group = interaction(posted_month, covid_preprint))) +
  geom_jitter(aes(fill = factor(covid_preprint)),
              shape = 21, size = 0.25, alpha = 0.1, 
              position = position_jitterdodge(jitter.width = 15, dodge.width = 21)) +
  geom_boxplot(width = 20, outlier.shape = NA, fill = NA,
               position = position_dodge(width = 21)) +
  labs(x = "Posted Month", y = "Total abstract views", 
       fill = "", color = "") +
  scale_y_log10(labels = scales::comma, limits = c(1, 1e6), expand = c(0, 0)) +
  scale_color_manual(values = qualitative_hcl(2, palette = "Set2")) +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  scale_x_date(date_breaks = "1 month",
               date_minor_breaks = "1 month",
               date_labels = "%b") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggsave("outputs/figures/partials/Figure_3A.png", width = 10, height = 4)

# PDF downloads
p3B <- d %>%
  ggplot(aes(x = posted_month, y = pdf_downloads, color = covid_preprint,
             group = interaction(posted_month, covid_preprint))) +
  geom_jitter(aes(fill = factor(covid_preprint)),
              shape = 21, size = 0.25, alpha = 0.1, 
              position = position_jitterdodge(jitter.width = 15, dodge.width = 21)) +
  geom_boxplot(width = 20, outlier.shape = NA, fill = NA,
               position = position_dodge(width = 21)) +
  labs(x = "Posted Month", y = "Total PDF downloads", 
       fill = "", color = "") +
  scale_y_log10(labels = scales::comma, limits = c(1, 1e6), expand = c(0, 0)) +
  scale_color_manual(values = qualitative_hcl(2, palette = "Set2")) +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  scale_x_date(date_breaks = "1 month",
               date_minor_breaks = "1 month",
               date_labels = "%b") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggsave("outputs/figures/partials/Figure_3B.png", width = 10, height = 4)

# Patchwork
p3A + p3B + plot_layout(ncol = 1) +
  plot_annotation(tag_levels = "A") +
  ggsave("outputs/figures/Figure_3.png", width = 10, height = 8)

```

# Supplementary Figure 3: Time prior to our study period
```{r}
# Panel A: Abstract views (bioRxiv + medRxiv, non-COVID)

d <- preprint_usage %>%
  inner_join(preprints, by = "doi") %>%
  filter(covid_preprint != T) %>%
  group_by(doi, posted_date) %>%
  summarize(full_text_views = sum(full_text_views),
            abstract_views = sum(abstract_views),
            pdf_downloads = sum(pdf_downloads)) %>%
  ungroup() %>%
  mutate(posted_month = floor_date(posted_date, unit = "month")) %>%
  filter(posted_date >= "2019-01-01",
         posted_date <= analysis_end)

pS3A <- d %>%
  ggplot(aes(x = posted_month, y = abstract_views, group = posted_month)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(shape = 21, size = 0.2, alpha = 0.1) +
  annotate("rect", xmin = ymd("2019-12-15"), xmax = ymd("2020-10-15"),
           ymin = 1, ymax = 1e5, fill = "red", alpha = 0.1) +
  labs(x = "Posted Month", y = "Total abstract views", 
       fill = "", color = "") +
  scale_y_log10(labels = scales::comma, limits = c(1, 1e5), expand = c(0, 0)) +
  scale_x_date(date_breaks = "1 month",
               date_minor_breaks = "1 month",
               date_labels = "%b-%y",
               expand = c(0.01,0.01)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggsave("outputs/figures/partials/Figure_S3A.png", width = 10, height = 4)

# Panel B: PDF downloads (bioRxiv + medRxiv, non-COVID)

pS3B <- d %>%
  ggplot(aes(x = posted_month, y = pdf_downloads, group = posted_month)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(shape = 21, size = 0.2, alpha = 0.1) +
  annotate("rect", xmin = ymd("2019-12-15"), xmax = ymd("2020-10-15"),
           ymin = 1, ymax = 1e5, fill = "red", alpha = 0.1) +
  labs(x = "Posted Month", y = "Total PDF downloads", 
       fill = "", color = "") +
  scale_y_log10(labels = scales::comma, limits = c(1, 1e5), expand = c(0, 0)) +
  scale_x_date(date_breaks = "1 month",
               date_minor_breaks = "1 month",
               date_labels = "%b-%y",
               expand = c(0.01,0.01)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggsave("outputs/figures/partials/Figure_S3B.png", width = 10, height = 4)

# Patchwork
pS3A + pS3B +
  plot_layout(nrow = 2) +
  plot_annotation(tag_levels = "A") +
  ggsave("outputs/figures/Figure_S3.png", width = 10, height = 8)

```


```{r}

# Will need to discuss if we add this back to Fig S3 or not

# Panel C: PDF downloads for additional preprint servers

# Bind into one data frame
all_server_dloads <- bind_rows(preprint_usage %>%
                                 inner_join(preprints, by = c("doi", "source")) %>%
                                 group_by(source, doi, posted_date, covid_preprint) %>%
                                 summarize(pdf_downloads = sum(pdf_downloads)) %>%
                                 ungroup() %>%
                                 select(-doi) %>%
                                 filter(posted_date >= analysis_start & posted_date <= analysis_end) ,
                               other_server_dloads)

# Generate plot
pS3C <- all_server_dloads %>%
  assign_covid_preprint() %>%
  ggplot(aes(x = source, y = pdf_downloads, color = covid_preprint)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(fill = factor(covid_preprint)),
              shape = 21, size = 0.2, alpha = 0.2,
              position = position_jitterdodge(jitter.width = 0.5)) + 
  labs(x = "Preprint server", y = "Total PDF downloads", 
       fill = "", color = "") +
  scale_y_log10(labels = scales::comma, limits = c(1, 1e6), expand = c(0, 0)) +
  scale_color_manual(values = qualitative_hcl(2, palette = "Set2")) +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  theme(legend.position = "top") + 
  ggsave("outputs/figures/partials/Figure_S3C.png", width = 10, height = 4)

# Patchwork
pS3A + pS3B + pS3C +
  plot_layout(nrow = 3) +
  plot_annotation(tag_levels = "A") +
  ggsave("outputs/figures/Figure_S3.png", width = 10, height = 10)

```

# Figure 4: Preprint usage and sharing

```{r}

# Prepare data
d <- preprints %>%
  filter(posted_date >= "2020-01-01") %>%
  inner_join(preprint_citations, by = "doi") %>%
  inner_join(preprint_comments %>% rename(comments = comments_count), by = "doi") %>%
  inner_join(preprint_altmetrics, by = "doi") %>%
  assign_covid_preprint() %>%
  mutate(posted_week = floor_date(posted_date, unit = "week", week_start = 1))

# Panel A: Citations per preprint (COVID vs non-COVID)

p4A <- d %>%
  ggplot(aes(x = covid_preprint, y = citations + 1, color = covid_preprint)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(fill = factor(covid_preprint)), shape = 21, size = 0.5, alpha = 0.4, width = 0.25) +
  labs(x = "", y = "Citations", 
       fill = "", color = "") +
  scale_y_log10(labels = scales::comma, expand = c(0, 0.1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = qualitative_hcl(2, palette = "Set2")) +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 12)) +
  guides(color = FALSE, fill = FALSE) +
  ggsave("outputs/figures/partials/Figure_4A.png", width = 5, height = 4)

# Panel B: Tweets per preprint (COVID vs non-COVID)

p4B <- d %>%
  ggplot(aes(x = covid_preprint, y = twitter + 1, color = covid_preprint)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(fill = factor(covid_preprint)), shape = 21, size = 0.5, alpha = 0.4, width = 0.25) +
  labs(x = "", y = "Tweets", 
       fill = "", color = "") +
  scale_y_log10(labels = scales::comma, expand = c(0, 0.1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = qualitative_hcl(2, palette = "Set2")) +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 12)) +
  guides(color = FALSE, fill = FALSE) +
  ggsave("outputs/figures/partials/Figure_4B.png", width = 5, height = 4)

# Panel C: News mentions per preprint (COVID vs non-COVID)

p4C <- d %>%
  ggplot(aes(x = covid_preprint, y = news + 1, color = covid_preprint)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(fill = factor(covid_preprint)), shape = 21, size = 0.5, alpha = 0.4, width = 0.25) +
  labs(x = "", y = "News Articles", 
       fill = "", color = "") +
  scale_y_log10(labels = scales::comma, expand = c(0, 0.1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = qualitative_hcl(2, palette = "Set2")) +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 12)) +
  guides(color = FALSE, fill = FALSE) +
  ggsave("outputs/figures/partials/Figure_4C.png", width = 5, height = 4)  

# Panel D: Blog mentions per preprint (COVID vs non-COVID)

p4D <- d %>%
  ggplot(aes(x = covid_preprint, y = blogs + 1, color = covid_preprint)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(fill = factor(covid_preprint)), shape = 21, size = 0.5, alpha = 0.4, width = 0.25) +
  labs(x = "", y = "Blogs", 
       fill = "", color = "") +
  scale_y_log10(labels = scales::comma, expand = c(0, 0.1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = qualitative_hcl(2, palette = "Set2")) +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 12)) +
  guides(color = FALSE, fill = FALSE) +
  ggsave("outputs/figures/partials/Figure_4D.png", width = 5, height = 4)

# Panel E: Count of comments per preprint (COVID vs non-COVID)
p4E <- d %>% 
  ggplot(aes(x = covid_preprint, y = comments + 1, color = covid_preprint)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(fill = factor(covid_preprint)), shape = 21, size = 0.5, alpha = 0.4, width = 0.25) +
  labs(x = "", y = "Comments", 
       fill = "", color = "") +
  scale_y_log10(labels = scales::comma, expand = c(0, 0.1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = qualitative_hcl(2, palette = "Set2")) +
  scale_fill_manual(values = qualitative_hcl(2, palette = "Set2")) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 12)) +
  guides(color = FALSE, fill = FALSE) +
  ggsave("outputs/figures/partials/Figure_4E.png", width = 5, height = 4) 

# Panels F and G: Correlation plots between citations, tweets, news articles, blogs and comments

covid_usage <- d %>% 
  filter(covid_preprint == "COVID-19 preprints") %>%
  select(citations, twitter, news, blogs, comments)

non_covid_usage <- d %>% 
  filter(covid_preprint == "non-COVID-19 preprints") %>%
  select(citations, twitter, news, blogs, comments)

covid_corr <- correlate(covid_usage, method = "spearman") %>%
  pivot_longer(citations:comments) %>%
  rename(rows = rowname,
         cols = name) %>%
  filter(str_trunc(rows, 2, ellipsis = "") < str_trunc(cols, 2, ellipsis = ""))

non_covid_corr <- correlate(non_covid_usage, method = "spearman") %>%
  pivot_longer(citations:comments) %>%
  rename(rows = rowname,
         cols = name) %>%
  filter(str_trunc(rows, 2, ellipsis = "") < str_trunc(cols, 2, ellipsis = ""))

p4F <- covid_corr %>%
  ggplot(aes(x = rows, y = cols, fill = value)) +
  geom_tile() +
  geom_text(aes(label = round(value, 2))) +
  scale_fill_continuous_sequential(palette = "Reds3", 
                                   breaks = c(0, 0.5, 1), 
                                   limits = c(0, 1)) +
  labs(x = "", y = "", title = "COVID-19 preprints",
       fill = "Spearman Correlation") +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 10)) +
  guides(fill = guide_colourbar(title.position="top", 
                                title.hjust = 0.5,
                                label.position = "bottom",
                                barwidth = 7)) +
  ggsave("outputs/figures/partials/Figure_4F.png", width = 5, height = 4)

p4G <- non_covid_corr %>%
  ggplot(aes(x = rows, y = cols, fill = value)) +
  geom_tile() +
  geom_text(aes(label = round(value, 2))) +
  scale_fill_continuous_sequential(palette = "Reds3", 
                                   breaks = c(0, 0.5, 1), 
                                   limits = c(0, 1)) +
  labs(x = "", y = "", title = "non-COVID-19 preprints",
       fill = "Spearman Correlation") +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 10)) +
  guides(fill = guide_colourbar(title.position="top", 
                                title.hjust = 0.5,
                                label.position = "bottom",
                                barwidth = 7)) +
  ggsave("outputs/figures/partials/Figure_4G.png", width = 5, height = 4)

# Panels H: Dotplot of proportion of citations that are preprints, per policy doc
p4H <- policy_docs %>%
#  filter(type == "COVID-19") %>%
  mutate(preprint_prop = npreprints*100/nrefs,
         set = case_when(set == "ECDC COVID-19 documents" ~ "ECDC",
                         set == "National academy of medicine" ~ "NAM",
                         set == "US committee COVID" ~ "US HSSCC",
                         set == "WHO Scientific Briefs" ~ "WHO SB",
                         set == "UK post" ~ "UK POST")) %>%
  ggplot(aes(x = set, y = preprint_prop)) +
  geom_dotplot(fill = qualitative_hcl(1, palette = "Set2"), color = qualitative_hcl(1, palette = "Set2"), binaxis = "y", binwidth = 1,  stackdir = "center", binpositions="all") +
  ylab("% preprints among citations") +
  guides(size = guide_legend(title="total\ncitations")) +
  theme(axis.title.x=element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggsave("outputs/figures/partials/Figure_4H.png", width = 4, height = 3)

# Patchwork
# Manually define layout
layout <- "
ABCD
#EF#
#GH#
#IJ#
"
p <- plot_spacer() + p4A + p4B + plot_spacer() + p4C + p4D + p4E + p4H + p4F + p4G + 
  plot_layout(design = layout, guides = "collect") & theme(plot.margin = margin(0, 0, 0, 0),
                                                           legend.title = element_text(size = 10),
                                                           legend.position = "bottom")

p +
  plot_annotation(tag_levels = "A") +
  ggsave("outputs/figures/Figure_4.png", width = 10, height = 12)

```

# Supplementary Figure 4: Twitter word clouds, sentiment

```{r}

# Here we generate Supplementary Figure 4. The method is a little messy, as we
# generate wordclouds and venn diagrams which do not seem to play well with 
# ggplot objects. Thus, we create each panel and save it as a separate PNG.
# Panels are then arranged using cowplot's "ggdraw" function. 

# Panel A: Word cloud

library(wordcloud2)
# The below packages are necessary for exporting wordclouds to PNG, see
# https://www.r-graph-gallery.com/196-the-wordcloud2-library.html
library("htmlwidgets")
library("webshot")
#webshot::install_phantomjs()
library(eulerr)
library(cowplot)

# Read tweet data (not shared publicly at the moment)
tweets <- read_csv("data/preprint_tweets_20190901_20200430.csv")

# First we make a wordcloud for tweets of all top-10 tweeted articles
top_preprints <- preprints %>% 
  inner_join(preprint_altmetrics, by = "doi") %>%
  filter(covid_preprint == T) %>% 
  top_n(10, twitter) %>% 
  arrange(desc(twitter)) %>% 
  pull(doi)

# Find hashtags associated with top-10 most tweeted covid-19 preprints, only original tweets
hashtags <- tweets %>%
  filter(doi %in% top_preprints) %>% 
  filter(is_retweet == F) %>% 
  select(hashtags) %>%
  na.omit() %>%
  mutate(hashtags = strsplit(as.character(hashtags), ",")) %>% 
  unnest(hashtags) %>%
  mutate(hashtags = str_remove_all(hashtags, "[^a-zA-Z0-9]"), # remove special characters
         hashtags = str_to_lower(hashtags)) %>% # convert all hashtags to lowercase
  filter(!hashtags %in% c("", "covid19", "coronavirus", "ncov2019", "covid", "covid2019", "sarscov2", "2019ncov", "hcov19", "19", "novelcoronavirus", "corona", "coronaovirus", "coronarovirus", "coronarvirus")) %>% # remove empty hashtags and specific keywords
  count(hashtags) %>%
  filter(n > 5) %>%
  mutate(n = log(n+1))

# Generate the wordcloud
p <- wordcloud2(hashtags, color = "random-dark", size = .25)

# Save the wordcloud to a PNG
saveWidget(p, "tweet_wordcloud.html", selfcontained = F)
webshot("tweet_wordcloud.html","outputs/figures/partials/Figure_S4A.png", 
        delay = 30, vwidth = 960, vheight = 720)

# Panel B: Euler diagram of overlap between the top-10 most tweeted, most mentioned
# in news and blogs, most cited and most commented. Here we use the eulerr package

top_twitter <- preprints %>%
  inner_join(preprint_altmetrics, by = "doi") %>%
  filter(covid_preprint == T) %>%
  top_n(10, twitter) %>%
  pull(doi)

top_news <- preprints %>%
  inner_join(preprint_altmetrics, by = "doi") %>%
  filter(covid_preprint == T) %>%
  top_n(10, news) %>%
  pull(doi)

top_blogs <- preprints %>%
  inner_join(preprint_altmetrics, by = "doi") %>%
  filter(covid_preprint == T) %>%
  top_n(10, blogs) %>%
  pull(doi)

top_cited <- preprints %>%
  inner_join(preprint_citations, by = "doi") %>%
  filter(covid_preprint == T) %>%
  top_n(10, citations) %>%
  pull(doi)

top_comments <- preprints %>%
  inner_join(preprint_comments, by = "doi") %>%
  filter(covid_preprint == T) %>%
  top_n(10, comments_count) %>%
  pull(doi)

fit <- list(
  Twitter = top_twitter,
  News = top_news,
  Blogs = top_blogs,
  Citations = top_cited,
  Comments = top_comments)

# Output the Venn diagram to a PNG
png(filename = "outputs/figures/partials/Figure_S4B.png",
    width = 960, height = 720)
plot(euler(fit), 
     fills = qualitative_hcl(n = 5, palette = "Set2"),
     alpha = 0.75,
     labels = list(cex = 1),
     quantities = list(cex = 1),
     width = 200)
dev.off()

# Now we fit all the PNG files together. Some manual adjustments may be necessary

pS4A <- ggdraw() +
  draw_image("outputs/figures/partials/Figure_S4A.png") +
  theme(plot.margin = margin(5, 5, 5, 5))

pS4B <- ggdraw() +
  draw_image("outputs/figures/partials/Figure_S4B.png") +
  theme(plot.margin = margin(5, 5, 5, 5))

# Patchwork

layout <- "
A
B
"

pS4A + pS4B +
  plot_layout(design = layout) +
  plot_annotation(tag_levels = "A") +
  ggsave("outputs/figures/Figure_S4.png",
         height = 10, width = 10)

```